\documentclass[a4paper]{article}

\usepackage{setup}

\hypersetup{pdfstartview=XYZ}%         zoom par defaut

\setlength{\droptitle}{-5em}   % This is your set screw
\title{\vspace{1.5cm}Some difficulties...}
\author{}
\date{\vspace{-5ex}}

\pagenumbering{gobble}

\theoremstyle{definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{property}{Property}
\newtheorem{corollary}{Corollary}

\begin{document}

  \maketitle

  \newcommand{\id} {\textsf{id}}
  
    \section{Utility of $\tenv$}

    Note: the use of $\tenv$ and the use of a fixpoint come together.
    Using $\tenv$ is useless if we don't use a fixpoint with it.
    Indeed, all the examples below can be rewritten with only one test, using pairs:
    \begin{align*}
      &\ite {e_1} {t_1} {(\ite {e_2} {t_2} {e} {\bot})} {\bot}\\
      \to& \ite {(e_1,e_2)} {(\pair {t_1} {t_2})} {e} {\bot}
    \end{align*}

    \subsection{Subject reduction}

    \begin{align*}
      &\ite {(x,\id y)} {\pair \Any \Any}
      {
        (\ite x \Int {\bpi_1 (x,y)} 0)
      } 0\\
      \leadsto
      &\ite {(x,y)} {\pair \Any \Any}
      {
        (\ite x \Int {\bpi_1 (x,y)} 0)
      } 0
    \end{align*}

    Explanation (without $\tenv$):
    \begin{itemize}
      \item The first expression has type $\Int$.
      \item The second expression has type $\Any$, as $(x,y) \in \dom\Gamma$ in the then branch.
    \end{itemize}

    Possible fix (without $\tenv$):
    \begin{itemize}
      \item Non-algorithmic typing rules such that we can choose to look at $\Gamma$ or not.\\
      We would still need type schemes.\\
      We need to write this new type system in an algorithmic way,
      otherwise we would still need an independent deterministic typeof (it would only move the problem).
      \item Somehow "repair" $\Gamma$ after/during refinements without needing $\tenv$ (just with the informations in $\Gamma$).
      It would need to check all elements in $\Gamma$ and strenghten them if needed.
      Would not be well-founded in the case of a ite statement in a test. So it seems more complicated and it does not solve any problem.
    \end{itemize}

    \subsection{Precision}

    \begin{align*}
      &\ite {(x,y)} {(\pair \Int \Int) \vee (\pair \Bool \Bool)}
      {
        (\ite x \Int y 0)
      } 0
    \end{align*}

    Explanation (without $\tenv$):
    \begin{itemize}
      \item This expression has type $\Int \vee \Bool$, altough we could expect it to be $\Int$.
      The reason for that is that relation between $x$ and $y$ after the first test (in the then branch) is lost,
      because we never give importance again to the fact that $(x,y)\in(\pair \Int \Int)\vee(\pair \Bool \Bool)$ during refinements.
    \end{itemize}

    Possible fix (without $\tenv$):
    \begin{itemize}
      \item ???
    \end{itemize}
    
    \section{Why we can't handle nested if statements precisely with $\tenv$}

    \begin{align*}
      \ite {((x,y), &\ite x \Int y 0)}
      {\pair {((\pair \Int \Int) \vee (\pair \Bool \Bool))} {(\Int \vee \Bool)}\\}
      {&(\ite x \Int y 0)\\} {&0}
    \end{align*}

    Explanation:
    \begin{itemize}
      \item If, during the refinement of the first test, we try to type the nested $\texttt{if}$ using the new $\tenv$ (the one that contains the current test itself),
      it would not be well-founded.
      \item If, at the opposite, we type the nested if with the old $\tenv$, then:
      \begin{enumerate}
        \item During the refinement, the nested $\texttt{if}$ will be typed without knowing that $(x,y)\in(\pair \Int \Int)\vee(\pair \Bool \Bool)$.
        So, its type will be equivalent to $\Gamma(y) \vee \Int$.
        \item In the same time, $\Gamma(y)$ will be refined to $\Int \vee \Bool$.
        \item In the end (after the fixpoint), the nested $\texttt{if}$ will be associated to the type $\Int \vee \Bool$.
        \item Now, if we try to retype the $\texttt{if}$ in the then branch (by forgetting the value that we computed earlier),
        we obtain the type $\Int$ (so a strictly stronger type).
        That's because, this time, $(x,y)\in(\pair \Int \Int)\vee(\pair \Bool \Bool)$ is in $\tenv$.
      \end{enumerate}
    \end{itemize}
    
\end{document}