\documentclass[a4paper]{article}

\usepackage{setup}

\hypersetup{pdfstartview=XYZ}%         zoom par defaut

\setlength{\droptitle}{-5em}   % This is your set screw
\title{\vspace{1.5cm}Type inference - M2 Internship}
\author{Mickael LAURENT}
\date{\vspace{-5ex}}

\pagenumbering{gobble}

\theoremstyle{definition}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{property}{Property}
\newtheorem{corollary}{Corollary}

\begin{document}

  \maketitle

  \subsubsection{Completeness}

    \begin{definition}[Bottom environment]
      Let $\Gamma$ an environment.\\
      $\Gamma$ is bottom (noted $\Gamma = \bot$) iff $\exists e\in\dom\Gamma.\ \Gamma(e)\simeq\Empty$.
    \end{definition}

    \begin{definition}[Algorithmic (pre)order on environments]
    Let $\Gamma$ and $\Gamma'$ two environments. We write $\Gamma' \leqA \Gamma$ iff:
    \begin{align*}
        &\Gamma'=\bot \text{ or } (\Gamma\neq\bot \text{ and } \forall e \in \dom \Gamma.\ \tyof e \Gamma \leq \Gamma(e))
    \end{align*}

    For an expression $e$, we write $\Gamma' \leqA^e \Gamma$ iff:
    \begin{align*}
      &\Gamma'=\bot \text{ or } (\Gamma\neq\bot \text{ and } \forall e' \in \dom \Gamma \text{ such that $e'$ is a subexpression of $e$}.\ \tyof {e'} \Gamma \leq \Gamma({e'}))
    \end{align*}

    Note that if $\Gamma' \leqA \Gamma$, then $\Gamma' \leqA^e \Gamma$ for any $e$.
    \end{definition}

    \begin{definition}[Order relation for type schemes]
      Let $\ts_1$ and $\ts_2$ two type schemes. We write $\ts_2 \leq \ts_1$ iff $\tsint {\ts_1} \subseteq \tsint{\ts_2}$.
    \end{definition}

    \begin{definition}[Positive derivation]
      A derivation of the declarative type system is said positive iff it does not contain any rule \Rule{Abs-}.
    \end{definition}

    \begin{definition}[Rank-0 negated derivation]
      A derivation of the declarative type system is said rank-0 negated iff any application of \Rule{PAppL}
      has a positive derivation as first premise ($\pvdash \Gamma e t \varpi.1:t_1$).
    \end{definition}

  \begin{lemma}
    \begin{align*}
      &\forall t,\ts.\ \tsrep{t\tsand\ts} \leq t \land \tsrep{\ts}\\
      &\forall \ts_1,\ts_2.\ \tsrep{\apply {\ts_1}{\ts_2}} \leq \apply {\tsrep {\ts_1}}{\tsrep {\ts_2}}
    \end{align*}
  \end{lemma}

  \begin{proof}
    Straightfoward, by induction on the structure of $\ts$.
  \end{proof}

  \begin{lemma}[Monotonicity of the algorithm] Let $\Gamma$, $\Gamma'$ and $e$ such that $\Gamma'\leqA^e \Gamma$ and $\tyof e \Gamma \neq \tsempty$. We have:
    \begin{align*}
      &\tyof e {\Gamma'} \leq \tyof e {\Gamma} \text{ and } \tsrep{\tyof e {\Gamma'}} \leq \tsrep{\tyof e {\Gamma}}\\
      &\forall t,\varpi.\ \env {\Gamma',e,t} (\varpi) \leq \env {\Gamma,e,t} (\varpi)\\
      &\forall t.\ \Refine {e,t} {\Gamma'} \leqA^e \Refine {e,t} \Gamma\\
    \end{align*}
  \end{lemma}

  \begin{proof} 
  We proceed by induction over the structure of $e$
  and, for two identical $e$, on the domains of $\Gamma$ and $\Gamma'$ (with the lexicographical inclusion order).

  Let's prove the first property: $\tyof e {\Gamma'} \leq \tyof e {\Gamma} \text{ and } \tsrep{\tyof e {\Gamma'}} \leq \tsrep{\tyof e {\Gamma}}$.
  We will focus on showing $\tyof e {\Gamma'} \leq \tyof e {\Gamma}$. The property $\tsrep{\tyof e {\Gamma'}} \leq \tsrep{\tyof e {\Gamma}}$
  can be proved in a very similar way, by using the fact that operators on type schemes like $\tsand$ or $\apply {} {}$ are also monotone.
  (Note that the only rule that introduces the type scheme constructor $\tsfun {\_}$ is \Rule{Abs\Aa}.)

  If $\Gamma' = \bot$ we can conclude directly with the rule \Rule{Efq}.
  So let's assume $\Gamma' \neq \bot$ and $\Gamma \neq \bot$
  (as $\Gamma = \bot \Rightarrow \Gamma' = \bot$ by definition of $\leqA^e$).

  If $e=x$ is a variable, then the last rule used in $\tyof e \Gamma$ and $\tyof e {\Gamma'}$ is \Rule{Var\Aa}.
  As $\Gamma' \leqA^e \Gamma$, we have $\Gamma'(e) \leq \Gamma(e)$ and thus
  we can conclude with the rule \Rule{Var\Aa}.
  So let's assume that $e$ is not a variable.

  If $e\in\dom\Gamma$, then the last rule used in $\tyof e \Gamma$ is \Rule{Env\Aa}.
  As $\Gamma' \leqA^e \Gamma$, we have $\tyof e {\Gamma'} \leq \Gamma(e)$.
  Moreover, by applying the induction hypothesis, we get $\tyof e {\Gamma'\setminus \{e\}} \leq \tyof e {\Gamma\setminus \{e\}}$
  (we can easily verify that $\Gamma'\setminus\{e\} \leqA^e \Gamma\setminus\{e\}$).
  \begin{itemize}
    \item If we have $e\in\dom{\Gamma'}$, we have according to the rule \Rule{Env\Aa}
    $\tyof e {\Gamma'} \leq \tyof e {\Gamma'\setminus \{e\}} \leq \tyof e {\Gamma\setminus \{e\}}$.
    Together with $\tyof e {\Gamma'} \leq \Gamma(e)$,
    we deduce $\tyof e {\Gamma'} \leq \Gamma(e) \tsand \tyof e {\Gamma\setminus \{e\}} = \tyof e {\Gamma}$.
    \item  Otherwise, we have $e\not\in\dom{\Gamma'}$. Thus
    $\tyof e {\Gamma'} = \tyof e {\Gamma'\setminus \{e\}} \leq \Gamma(e) \tsand \tyof e {\Gamma\setminus \{e\}}=\tyof e {\Gamma}$.
  \end{itemize}

  If $e\not\in\dom\Gamma$ and $e\in\dom{\Gamma'}$, the last rule is \Rule{Env\Aa} for $\tyof e {\Gamma'}$.
  As $\Gamma'\setminus\{e\} \leqA^e \Gamma\setminus\{e\} = \Gamma$,
  we have $\tyof e {\Gamma'} \leq \tyof e {\Gamma'\setminus \{e\}} \leq \tyof e {\Gamma}$ by induction hypothesis.

  Thus, let's suppose that $e\not\in\dom\Gamma$ and $e\not\in\dom{\Gamma'}$.
  From now we know that the last rule in $\tyof e {\Gamma}$ and $\tyof e {\Gamma'}$ is the same.

  \begin{description}
    \item[$e=c$] The last rule is \Rule{Const\Aa}. It does not depend on $\Gamma$ so this case is trivial.
    \item[$e=x$] Already treated.
    \item[$e=\lambda^{\bigwedge_{i\in I} \arrow {t_i}{s_i}}x.e'$]
    The last rule is \Rule{Abs\Aa}.
    We have $\forall i\in I.\ \Gamma', (x:s_i) \leqA^{e'} \Gamma, (x:s_i)$ (quite straightforward)
    so by applying the induction hypothesis we have $\forall i\in I.\ \tyof {e'} {\Gamma', (x:s_i)} \leq \tyof {e'} {\Gamma, (x:s_i)}$.

    \item[$e=e_1 e_2$] The last rule is \Rule{App\Aa}.
    We can conclude immediately by using the induction hypothesis and noticing that $\apply {} {}$ is monotonic for both of its arguments.

    \item[$e=\pi_i e'$] The last rule is \Rule{Proj\Aa}.
    We can conclude immediately by using the induction hypothesis and noticing that $\bpi_i$ is monotonic.
     
    \item[$e=(e_1,e_2)$] The last rule is \Rule{Pair\Aa}.
    We can conclude immediately by using the induction hypothesis.

    \item[$\ite {e_0} t {e_1} {e_2}$] The last rule is \Rule{Case\Aa}.
    By using the induction hypothesis we get $\Refine {e_0,t} {\Gamma'} \leqA^{e_0} \Refine {e_0,t} \Gamma$.
    We also have $\Gamma' \leqA^{e_1} \Gamma$ (as $e_1$ is a subexpression of $e$).

    From those two properties, let's show that we can deduce $\Refine {e_0,t} {\Gamma'} \leqA^{e_1} \Refine {e_0,t} \Gamma$:

    Let $e'\in\dom{\Refine {e_0,t} \Gamma}$ a subexpression of $e_1$.
    \begin{itemize}
      \item If $e'$ is also a subexpression of $e_0$, we can directly deduce
      $\tyof {e'} {\Refine {e_0,t} \Gamma'} \leq (\Refine {e_0,t} \Gamma) (e')$
      by using $\Refine {e_0,t} {\Gamma'} \leqA^{e_0} \Refine {e_0,t} \Gamma$.

      \item Otherwise, as $\Refine {e_0,t} {\_}$ is reductive,
      we have $\Refine {e_0,t} {\Gamma'} \leqA \Gamma'$ and thus by using the induction hypothesis
      $\tyof {e'} {\Refine {e_0,t} {\Gamma'}} \leq \tyof {e'} {\Gamma'}$.
      We also have $\tyof {e'} {\Gamma'} \leq \Gamma(e')$ by using $\Gamma' \leqA^{e_1} \Gamma$.
      We deduce $\tyof {e'} {\Refine {e_0,t} {\Gamma'}} \leq \Gamma(e') = (\Refine {e_0,t} \Gamma) (e')$.
    \end{itemize}
    
    So we have $\Refine {e_0,t} {\Gamma'} \leqA^{e_1} \Refine {e_0,t} \Gamma$.
    Consequently, we can apply the induction hypothesis again to get
    $\tyof {e_1} {\Refine {e_0,t} {\Gamma'}} \leq \tyof {e_1} {\Refine {e_0,t} {\Gamma}}$.

    We proceed the same way for the last premise.
  \end{description}\ \\

  Now, let's prove the second property.
  We perform a (nested) induction on $\varpi$.

  We know that $\tsrep{\tyof {\occ e \varpi} {\Gamma'}} \leq \tsrep{\tyof {\occ e \varpi} {\Gamma}}$
  by the outer induction hypothesis (for $\varpi=\epsilon$ we use the proof of the first property above).

  Thus we just have to prove that $\constr {\varpi} {\Gamma',e,t} \leq \constr {\varpi} {\Gamma,e,t}$.
  The only case that is interesting is the case $\varpi=\varpi'.1$.

  First, we can notice that the $\worra {} {}$ operator is monotonic for its second argument
  (consequence of its declarative definition).

  Secondly, let's show that for any function types $t_1 \leq t_2$, and for any type $t'$,
  we have $(\worra {t_1} {t'}) \land \dom {t_2} \leq \worra {t_2} {t'}$. By the absurd, let's suppose it is not true.
  Let's note $t'' = (\worra {t_1} {t'}) \land \dom {t_2}$.
  Then we have $t'' \leq \dom {t_2} \leq \dom {t_1}$ and $t_2 \leq \arrow {t''} {t'}$ and
  $t_1 \not\leq \arrow {t''} {t'}$, which contradicts $t_1 \leq t_2$.

  Let's note $t_1 = \tsrep {\tyof{\occ e{\varpi'.0}}{\Gamma'}}$ and $t_2 = \tsrep {\tyof{\occ e{\varpi'.0}}\Gamma}$
  and $t'=\env {\Gamma,e,t} (\varpi')$.
  As $e$ is well-typed, and using the inner induction hypothesis, we have $\tsrep {\tyof {\occ e {\varpi'.1}} {\Gamma'}} \leq \tsrep {\tyof {\occ e {\varpi'.1}} {\Gamma}} \leq \dom {t_2}$.\\
  Thus, using this property, we get:\\
  \begin{align*}
  &(\worra {t_1} {t'}) \land \tsrep {\tyof {\occ e {\varpi'.1}} {\Gamma'}}\\
  \leq &(\worra {t_2} {t'}) \land \tsrep {\tyof {\occ e {\varpi'.1}} {\Gamma}}
  \end{align*}

  Then, using the monotonicity of the second argument of $\worra {}{}$ and the outer induction hypothesis:
  \begin{align*}
    &(\worra {t_1} {\env {\Gamma',e,t} (\varpi')}) \land \tsrep {\tyof {\occ e {\varpi'.1}} {\Gamma'}}\\
    \leq &(\worra {t_2} {\env {\Gamma,e,t} (\varpi')}) \land \tsrep {\tyof {\occ e {\varpi'.1}} {\Gamma}}
  \end{align*}
  \\

  \ 

  Finally, we must prove the third property.\\
  It is straightforward by using the previous result and the induction hypothesis:\\
  for any $e'$ such that $\exists \varpi.\ \occ e \varpi \equiv e'$, we get
  $\bigwedge_{\{\varpi\alt \occ e \varpi \equiv e'\}} \env {\Gamma',e,t} (\varpi) \leq \bigwedge_{\{\varpi\alt \occ e \varpi \equiv e'\}} \env {\Gamma,e,t} (\varpi)$.

  The rest follows.
  \end{proof}

  \begin{theorem}[Completeness for positive derivations]
    For every $\Gamma$, $e$, $t$ such that we have a positive derivation of $\Gamma \vdash e:t$,
    there exists a global parameter $n_o$ with which $\tsrep{\tyof e \Gamma} \leq t$.

    More precisely:
    \begin{align*}
      &\forall \Gamma, e, t.\ \Gamma \vdash e:t \text{ has a positive derivation } \Rightarrow \tsrep{\tyof e \Gamma} \leq t\\
      &\forall \Gamma, \Gamma', e, t.\ \Gamma \evdash e t \Gamma' \text{ has a positive derivation } \Rightarrow \Refine {e,t} \Gamma \leqA \Gamma' \text{ (for $n_o$ large enough)}
    \end{align*}
  \end{theorem}

  \begin{proof}
  We proceed by induction on the derivation.

  Let's prove the first property. We have a positive derivation of $\Gamma \vdash e:t$.

  If $\Gamma = \bot$, we can conclude directly using \Rule{Efq\Aa}. Thus, let's suppose $\Gamma \neq \bot$.

  If $e=x$ is a variable, then the derivation only uses \Rule{Env}, \Rule{Inter} and \Rule{Subs}.
  We can easily conclude just be using \Rule{Var\Aa}. Thus, let's suppose $e$ is not a variable.

  If $e\in\dom\Gamma$, we can have the rule \Rule{Env} applied to $e$ in our derivation, but in this case
  there can only be \Rule{Inter} and \Rule{Subs} after it (not \Rule{Abs-} as we have a positive derivation).
  Thus, our derivation contains a derivation of $\Gamma \vdash e:t'$ that does not use the rule \Rule{Env} on $e$
  and such that $t'\land \Gamma(e) \leq t$ (actually, it is possible for our derivation to typecheck $e$ only using the rule \Rule{Env}:
  in this case we can take $t'=\Any$ and use the fact that $\Gamma$ is well-formed).
  Hence, we can build a positive derivation for $\Gamma\setminus\{e\} \vdash e:t'$.
  By using the induction hypothesis we deduce that $\tsrep{\tyof e {\Gamma\setminus\{e\}}} \leq t'$.
  Thus, by looking at the rule \Rule{Env\Aa},
  we deduce $\tsrep{\tyof e \Gamma} \leq \Gamma(e) \land \tsrep{\tyof e {\Gamma\setminus\{e\}}} \leq t$.
  It concludes this case, so let's assume $e\not\in\dom\Gamma$.

  Now we analyze the last rule of the derivation:

  \begin{description}
    \item[\Rule{Env}] Impossible case ($e\not\in\dom\Gamma$).
    \item[\Rule{Inter}] By using the induction hypothesis we get $\tsrep{\tyof e \Gamma} \leq t_1$ and $\tsrep{\tyof e \Gamma} \leq t_2$.
    Thus, we have $\tsrep{\tyof e \Gamma} \leq t_1 \land t_2$. 
    \item[\Rule{Subs}] Trivial using the induction hypothesis.
    \item[\Rule{Const}] We know that the derivation of $\tyof e \Gamma$ ends with the rule \Rule{Const\Aa}.
    Thus this case is trivial.
    \item[\Rule{App}] We know that the derivation of $\tyof e \Gamma$ ends with the rule \Rule{App\Aa}.
    Let $\ts_1 = \tyof {e_1} \Gamma$ and $\ts_2 = \tyof {e_2} \Gamma$. 
    With the induction hypothesis we have $\tsrep {\ts_1} \leq \arrow {t_1} {t_2}$ and $\tsrep {\ts_2} \leq t_1$, with $t_2=t$.
    According to the descriptive definition of $\apply{}{}$, we have
    $\apply{\tsrep {\ts_1}}{\tsrep {\ts_2}} \leq \apply{\arrow {t_1}{t_2}}{t_1} \leq t_2$.
    As we also have $\tsrep{\apply {\ts_1} {\ts_2}} \leq \apply{\tsrep {\ts_1}}{\tsrep {\ts_2}}$,
    we can conclude that $\tyof e \Gamma \leq t_2=t$.

    \item[\Rule{Abs+}] We know that the derivation of $\tyof e \Gamma$ ends with the rule \Rule{Abs\Aa}.
    This case is straightforward using the induction hypothesis.
    \item[\Rule{Abs-}] This case is impossible (the derivation is positive).
    \item[\Rule{Case}] We know that the derivation of $\tyof e \Gamma$ ends with the rule \Rule{Case\Aa}.
    By using the induction hypothesis and the monotonicity lemma, we get $\tsrep{\ts_1}\leq t$ and $\tsrep{\ts_2}\leq t$.
    So we have $\tsrep{\ts_1\tsor\ts_2}=\tsrep{\ts1}\vee\tsrep{\ts2}\leq t$.
    \item[\Rule{Proj}] Quite similar to the case \Rule{App}.
    \item[\Rule{Pair}] We know that the derivation of $\tyof e \Gamma$ ends with the rule \Rule{Pair\Aa}.
    We just use the induction hypothesis and the fact that $\tsrep{\ts_1\tstimes\ts_2}=\pair {\tsrep{\ts1}} {\tsrep{\ts2}}$.
  \end{description}

  \ 

  Now, let's prove the second property. We have a positive derivation of $\Gamma \evdash e t \Gamma'$.

  \begin{description}
    \item[\Rule{Base}] Any value of $n_o$ will give $\Refine {e,t} \Gamma \leqA \Gamma$, even $n_o = 0$.
    \item[\Rule{Path}] We have $\Gamma' = \Gamma_1,(\occ e \varpi:t')$.
    By applying the induction hypothesis on the premise $\Gamma \evdash e t \Gamma_1$, we have
    $\RefineStep {e,t}^n (\Gamma) = \Gamma_2$ with $\Gamma_2 \leqA \Gamma_1$ for a certain $n$.

    We now proceed by induction on the derivation $\pvdash {\Gamma_1} e t \varpi:t'$
    to show that we can obtain $\env {\Gamma'',e,t} (\varpi) \leq t'$ with $\RefineStep {e,t}^{n'} (\Gamma_2) \leqA \Gamma''$
    for a certain $n'$. It is then easy to conclude by taking $n_o = n+n'$.

    \begin{description}
      \item[\Rule{PSubs}] Trivial using the induction hypothesis.
      \item[\Rule{PInter}] By using the induction hypothesis we get
      $\env {\Gamma_1'',e,t} (\varpi) \leq t_1$ and $\env {\Gamma_2'',e,t} (\varpi) \leq t_2$
      with $\RefineStep {e,t}^{n_1} (\Gamma_2) \leqA \Gamma_1''$ and
      $\RefineStep {e,t}^{n_2} (\Gamma_2) \leqA \Gamma_2''$. By taking $n'=\max (n_1,n_2)$,
      we have $\RefineStep {e,t}^{n'} (\Gamma_2) \leqA \Gamma''$ with $\Gamma'' \leqA \Gamma_1''$ and $\Gamma'' \leqA \Gamma_2''$.
      Thus, by using the monotonicity lemma, we can obtain $\env {\Gamma'',e,t} (\varpi) \leq t_1 \land t_2 = t'$.
      \item[\Rule{PTypeof}] By using the outer induction hypothesis we get
      $\tsrep{\tyof {\occ e \varpi} {\Gamma_2}} \leq t'$.
      Moreover we have $\env {\Gamma_2,e,t} (\varpi) \leq \tsrep{\tyof {\occ e \varpi} {\Gamma_2}}$
      (by definition), thus we can conclude directly.
      \item[\Rule{PEps}] Trivial.
       
      \item[\Rule{PAppR}] By using the induction hypothesis we get
      $\env {\Gamma_1'',e,t} (\varpi.0) \leq \arrow {t_1} {t_2}$ and $\env {\Gamma_2'',e,t} (\varpi) \leq t_2'$
      with $\RefineStep {e,t}^{n_1} (\Gamma_2) \leqA \Gamma_1''$, $\RefineStep {e,t}^{n_2} (\Gamma_2) \leqA \Gamma_2''$
      and $t_2\land t_2' \simeq \Empty$.
      
      By taking $n'=\max (n_1,n_2) + 1$,
      we have $\RefineStep {e,t}^{n'} (\Gamma_2) \leqA \Gamma''$ with $\Gamma'' \leqA \RefineStep {e,t} (\Gamma_1'')$
      and $\Gamma'' \leqA \RefineStep {e,t} (\Gamma_2'')$.
      In consequence, we have $\tsrep{\tyof {\occ e {\varpi.0}} {\Gamma''}} \leq \env {\Gamma_1'',e,t} (\varpi.0) \leq \arrow {t_1} {t_2}$
      (by definition of $\RefineStep {e,t}$).
      We also have, by monotonicity, $\env {\Gamma'',e,t} (\varpi) \leq t_2'$.

      As $t_2\land t_2' \simeq \Empty$, we have:
      \begin{align*}
        &\apply {(\arrow {t_1} {t_2})} {(\dom{\arrow {t_1} {t_2}}\setminus(\neg t_1))}\\
        &\simeq \apply {(\arrow {t_1} {t_2})} {t_1} \simeq t_2 \leq \neg t_2'
      \end{align*}

      Thus, by using the declarative definition of $\worra {} {}$, we know that
      $\worra {(\arrow {t_1} {t_2})} {t_2'} \leq \neg t_1$.

      According to the properties on $\worra {} {}$ that we have proved in the proof of the monotonicity lemma,
      we can deduce:
      \begin{align*}
      &t_1 \land \worra {\tsrep{\tyof {\occ e {\varpi.0}} {\Gamma''}}} {\env {\Gamma'',e,t} (\varpi)}\\
      &\leq t_1 \land \worra {(\arrow {t_1} {t_2})} {t_2'} \leq t_1 \land \neg t_1 \simeq \Empty
      \end{align*}
      And thus $\worra {\tsrep{\tyof {\occ e {\varpi.0}} {\Gamma''}}} {\env {\Gamma'',e,t} (\varpi)} \leq \neg t_1$.

      It concludes this case.

      \item[\Rule{PAppL}] By using the induction hypothesis we get
      $\env {\Gamma_1'',e,t} (\varpi.1) \leq t_1$ and $\env {\Gamma_2'',e,t} (\varpi) \leq t_2$
      with $\RefineStep {e,t}^{n_1} (\Gamma_2) \leqA \Gamma_1''$ and
      $\RefineStep {e,t}^{n_2} (\Gamma_2) \leqA \Gamma_2''$. By taking $n'=\max (n_1,n_2)$,
      we have $\RefineStep {e,t}^{n'} (\Gamma_2) \leqA \Gamma''$ with $\Gamma'' \leqA \Gamma_1''$ and $\Gamma'' \leqA \Gamma_2''$.
      Thus, by using the monotonicity lemma, we can obtain $\env {\Gamma'',e,t} (\varpi.0) \leq \neg (\arrow {t_1} {\neg t_2}) = t'$.
  
      \item[\Rule{PPairL}] Trivial using the induction hypothesis and the descriptive definition of $\bpi_1$.
      \item[\Rule{PPairR}] Trivial using the induction hypothesis and the descriptive definition of $\bpi_2$.
      \item[\Rule{PFst}] Trivial using the induction hypothesis.
      \item[\Rule{PSnd}] Trivial using the induction hypothesis.
    \end{description}
  \end{description}
  \end{proof}

  \begin{corollary}\label{app:completeness}
    If we restrict our language so that the only functional type that can be used in a test is $\arrow \Empty \Any$
    and such that there is no negated arrows in the domain of lambda-abstractions,
    then the algorithmic type system becomes complete
    (i.e. any expression that can be typed in the declarative type system is also typeable in the algorithmic one)
    and type schemes can be removed from it (we can use regular types instead).
  \end{corollary}

  \begin{proof}
    With such restrictions, the rule \Rule{Abs-} is not needed anymore.

    Indeed, when typing an application $e_1 e_2$, the negative part of the type of $e_1$ (i.e. the $N_i$ part of its DNF)
    is ignored by the operator $\apply {} {}$.

    Moreover, as there is no negated arrows in the domain of lambda-abstractions,
    the negative arrows of the type of $e_2$ can also be ignored.

    Similarly, negative arrows can also be ignored when refining an application ($\worra {} {}$ also ignore the negative part
    of the type of $e_1$).

    Finally, as the only functional type that we can test is $\arrow \Empty \Any$, a functional type
    cannot be refined to $\Empty$ due to its negative part, and thus we can ignore its negative part
    (it makes no difference relatively to the rule \Rule{Efq}).
  \end{proof}

  \begin{theorem}[Completeness for rank-0 negated derivations]
    For every $\Gamma$, $e$, $t$ such that we have a rank-0 negated derivation of $\Gamma \vdash e:t$, there exists a global parameter $n_o$
    with which $\tyof e \Gamma \leq t$.

    More precisely:
    \begin{align*}
      &\forall \Gamma, e, t.\ \Gamma \vdash e:t \text{ has a rank-0 negated derivation } \Rightarrow \tyof e \Gamma \leq t\\
      &\forall \Gamma, \Gamma', e, t.\ \Gamma \evdash e t \Gamma' \text{ has a rank-0 negated derivation } \Rightarrow \Refine {e,t} \Gamma \leqA \Gamma' \text{ (for $n_o$ large enough)}
    \end{align*}
  \end{theorem}

  \begin{proof}
    This proof is quite similar to that of the completeness for positive derivations.
    In consequence, we will only detail cases that are quite different from those of the previous proof.

    Let's begin with the first property. We have a rank-0 negated derivation of $\Gamma \vdash e:t$.
    We want to show $\tyof e \Gamma \leq t$ (note that this is weaker than showing $\tsrep {\tyof e \Gamma} \leq t$).

    As in the previous proof, we can suppose that $\Gamma \neq \bot$ and that $e$ is not a variable.

    The case $e\in\dom\Gamma$ is also very similar, but there is an additional case to consider:
    the rule \Rule{Abs-} could possibly be used after a rule \Rule{Env} applied on $e$.
    However, this case can easily be eliminated by changing the premise of this \Rule{Abs-} with another one
    that does not use the rule \Rule{Env} on $e$ (the type of the premise does not matter for the rule \Rule{Abs-},
    even $\Any$ suffices).

    TODO
  \end{proof}

\end{document}